{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}

### 2014-08-22 swmp_noise.tex
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Mean correlation for all time series and window widths between the detided and biological values was 0.56, with values ranging from -0.78 to 1.00.  Mean error was 1.27, with values ranging from 0.00 to 2.47.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.  While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).  Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.  Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.  For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).

@@

Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to further validate the weighted regression model.  Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}). Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}). Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.  Each station was also chosen based on high correlations between \ac{DO} and tidal changes.  The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).  The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.  Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.

\subsection{Estimates of ecosystem metabolism before and after detiding}

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.  The half window widths that produced the most robust results for the simulations were used for the case studies: eight days, six hours, and a tidal proportion of one. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

@@

The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).  Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the mean annual values. Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (0.05 for production, 0.06 for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (0.11 for production, 0.12 for respiration).  Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.

Two examples from Rookery Bay and Sapelo Island illustrate the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1,fig:case_ex2}).  Tidal predictions from each site indicate that both are macrotidal, whereas Rookery Bay has more mixed semidiurnal tidal frequencies compared to more strongly semidiurnal frequencies at Sapelo Island. The effects of tidal height changes are apparent at both sites such that the observed \ac{DO} time series in the middle plots closely resemble the tidal height predictions on the lower plots.  Metabolism estimates based on the observed \ac{DO} time series were also largely influenced by tidal variation with anomalous esimates observed from August 11--15 at Rookery Bay (\cref{fig:case_ex1}) and February 4--7, 16, and 19 at Sapelo Island (\cref{fig:case_ex2}). These anomalous values occur when tidal height changes are out of phase with solar periods such that water masses with different spatial origins and metabolic histories are moving past the sensor during these periods.  Increasing tide during the night causes an influx of oxygen rich water from the ocean when a decrease in \ac{DO} from respiration is expected, whereas decreasing tide during the day causes an influx of anoxic water from landward areas when an increase in \ac{DO} from production is expected.  Similarly, a positive bias in production and negative bias in respiration would be expected when the tidal changes are in phase with solar periods.  Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide, which is further supported by more obvious diel variation with solar periods in the detided \ac{DO} time series.  Similar trends in metabolism estimates were observed for all sites across the period of observation.

@@

A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.  For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.  The effects of tidal variation on estimates of ecosystem metabolism may not be a concern mean annual rates are the primary focus.  Conversely, production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay, suggesting caution in interpreting averages of metabolism estimates on longer time periods if observed \ac{DO} data are not detided. Regardless, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.  For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated if aggregation of observed data on longer time periods removes potential bias from detiding?

The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate differences in mean metabolism estimates as a basis for evaluating aggregation effects on different periods of observation (\cref{fig:metab_sum1,fig:metab_sum2}).  Interestingly, estimates of net metabolism were similar regardless of the period of observation for which values were aggregated.  This suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at relatively short periods of observation (e.g., monthly), although note that substantial within-period variability may be observed compared to using a detided \ac{DO} time series.  Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.  Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.  For example, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data. Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.  Regardless of trends, these results suggest that tidal variation can affect interpretations of metabolic trends that are based on aggregations of daily values, particularly those as short as one month.

@@

Results from the simulations and case studies suggested that weighted regression is a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.  However, application of the method may only be appropriate under specific situations.  The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  The results for Elkhorn Slough were cause for concern given that the detided metabolism estimates were substantially reduced compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding was unusually large.  The detiding results are highly sensitive to changes in the window widths, particularly the hour window.  Further investigation of the effects of different window widths on the metabolism estimates for Elkhorn Slough were unable to produce satisfactory results such that the metabolism estimates were significantly lower in all cases.  This suggests that the weighted regression was removing variation in the observed \ac{DO} related to both tidal changes and metabolic components.

The results for Elkhorn Slough provide an example that emphasizes the need for restraint in applying weighted regression.  Because the true biological \ac{DO} signal is not known, no single measure can provide a true representation of the effectiveness of weighted regression.  Accordingly, we propose a precautionary approach that emphasizes application of the method following a set of conditions and steps.  First, weighted regression should only be applied at macrotidal sites and if there is strong evidence of the confounding effects of tidal advection.  The occurrence of anomalous estimates alone may not be sufficient for justifying use of the method.  A simple scatterplot of observed \ac{DO} versus tide height should indicate a strong correlation, in addition to clear visual patterns of tide change on \ac{DO}.  Second, significant changes in mean metabolic rates before and after detiding warrants a more comprehensive application of the method, with attention on specific window widths.  Changes in mean annual rates are of particular concern, which suggest that the method has reduced both physical and biological variation in the \ac{DO} signal.  Third, optimal results may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.  In general, a pragmatic approach is emphasized such that results based on detided estimates should be evaluated for the preservation of diel variation while exhibiting minimal changes with the tide.  Such an approach will support more accurate estimates of ecosystem metabolism that fulfill assumptions of the open-water method.

% caution when mean changes

@@

{\centering \includegraphics[width=\maxwidth]{figure/sim_ex}

@@

\caption[Representative examples of simulated \ac{DO} time series created by varying each of five parameters]{Representative examples of simulated \ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).  Black lines are observed \ac{DO} from \cref{do_obs_all} and red lines are biological \ac{DO} from \cref{do_bio}.\label{fig:sim_ex}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.  Results are averaged for all window widths used to evaluate the regressions.\label{fig:err_surf1}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.  Results are averaged for all simulation parameters used to evaluate the regressions.\label{fig:err_surf2}}
\end{figure}

@@

% example from RKBMB
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}

@@

\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression. Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot. Results are for the Rookery Bay station for twenty days in August, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
\end{figure}

@@

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}

@@

\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression. Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot. Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex2}}
\end{figure}

>>>>

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, and degree of association of the tide with the \ac{DO} signal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Three half window widths for each variable (day, hour of day, and tide height) were evaluated: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.  Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.  Mean error was 1.21, with values ranging from 0.00 to 2.40.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.  While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.  Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.  For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).

@@

Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the detiding model with real data.  Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}). Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}). Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.  Each station was also chosen based on high correlations between \ac{DO} and tidal changes.  The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).  The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.  Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.

\subsection{Estimates of ecosystem metabolism before and after detiding} \label{met_sec}

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.  Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although we recognize that the chosen widths are somewhat arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management actions. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

@@

The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).  Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the daily estimates. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).  Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.  Graphical results for each case study can be viewed using the link in the \hyperref[multi]{multimedia} section.

An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).   A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.  The first week illustrates a period of observation when maximum tide heights are generally out of phase with the diel cycle such that low tides are observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).  The second week illustrates when the maximum tide height occurs during the day and night (\cref{fig:phase_in}).  The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.  Maximum or minimum tidal heights also occur when the contributions of respiration and production are largest (i.e., mid-day or mid-night). The first example illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides, whereas the second example illustrates a strong positive bias (more respiraiton, more production) in the observed \ac{DO} from high tides. These biases are apparent in the metabolism estimates based on the observed data (\cref{fig:case_ex}).  Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.  The detided time series shows noticeable changes given the direction of bias associated with tidal height and diel period.  \ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods.  Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.  Detiding had similar effects for the remaining sites, particularly when tidal changes were in or out of phase with diel periods.

@@

A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.  The example from Sapelo Island in the previous section further highlights this point given that mean production and respiration estimates before and after detidng were generally unchanged for the two-week period. \cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay. Although we acknowledge that the specific results may be related to the window widths, this suggests that detiding may contribute to changes in metabolic estimates that are aggregated on longer time periods.  Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.  Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?  A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.

The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates (\cref{fig:metab_sum1,fig:metab_sum2}).  Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation. Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.  Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.  Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data. Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.  Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.  Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.  Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).  Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.

@@

Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.  However, application of the method may only be appropriate under specific situations.  The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are particular cause for concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.  In particular, optimal window widths should be chosen based on the ability to remove tidal effects while preserving the biological signal.

The case studies emphasize a critical challenge in evaluating the effects of physical advection: the true biological signal is not known, or more appropriately, the relative contribution of horizontal advection to bias is not accuratley quantified.  Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate performance of weighted regression.  Regardless, the current analysis suggests the method is potentially useful and we propose a precautionary approach for its application pending further validation.  Weighted regression should only be applied at macrotidal sites with strong evidence of the confounding effects of tidal advection on biological signals.  A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.  The choice of window widths may also produce varying results.  Optimal widths may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.  Results from window widths that produce large changes in the mean annual estimates should interpreted with caution.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation related to production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

@@

{\centering \includegraphics[width=\textwidth]{figure/sim_ex}

@@

\caption[Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters]{Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Parameter values represent the minimum and maximum used in the simulations.\label{fig:sim_ex}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.  Results are averaged for all window widths used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf1}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.  Results are averaged for all simulation parameters used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf2}}
\end{figure}

@@

% example from SAPHD, phase out
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/phase_out}

@@

\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$). Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during night and day periods. Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
\end{figure}

% example from SAPHD, phase in
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/phase_in}

\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$). Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during night and day periods. Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
\end{figure}

@@

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex}

@@

\caption[Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression. Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\cref{fig:phase_out}) and in phase during the second week (\cref{fig:phase_in}).\label{fig:case_ex}}
\end{figure}
